---
title: "Report"
output: html_document
---

**Flower classification project**
*Komari Machida and Indigo Morgenstern*

The purpose of this project is to develop a function in R that uses a bivariate normal distribution to classify a flower as either *iris versicolor* or *iris virginica* based on the width and length of its petals. For our training set, we use the built-in **iris** data package.

(A) To start, we import the package and create an initial plot to demonstrate the overall trend of the respective species' petal lengths and widths.

```{r}
data(iris)
d<-iris

library(tidyverse)

iris %>%
  filter(Species %in% c("versicolor", "virginica")) %>%
  ggplot(aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  geom_point()
```

(B) This initial graphic gives us a sense of how a human could classify a flower given the available data. For instance, if the flower has petal length of 4.0 and petal width of 1.1, then we can assume it is a *versicolor.* Likewise, if the length is 6 and its width is 2.3, we can assume it is a *virginica.*

However, if the length is 5.0 and width is 1.7, it is unclear which species it is most likely to be. Using a K-nearest-neighbor algorithm with K=3, it would be classified as a *versicolor.* However, with K=5, it would be classified as a *viriginica.*

(C) In the interest of a more rigorous mathematical classification model, we turn to Bayes' rule to derive a classification rule based on the probability of the flower being one type as opposed to another:

$$
\begin{aligned}
  P(Y=0|X=x) &> P(Y=1|X=x) \\
  \frac{p_0f_0(x)}{p_0f_0(x)+p_1f_1(x)} &> \frac{p_1f_1(x)}{p_0(x)f_0(x)+p_1(x)f_1(x)} \\
  p_0f_0(x) &> p_1f_1(x)
\end{aligned}
$$
[add explanations in derivation above? nahh]

(D) Because we don't know the exact values of the variables in the derived inequality, we must estimate them. To start, we can use the following unbiased estimator for $p_0$:

$$
\hat{p_0} = \frac{\text{number of flowers of type 0 in the training set}}{\text{total number of flowers in the training set}}
$$
We know it is unbiased because its bias is zero. This is because [explanation of how to get from line 1 to line 2: probability of the flower being one type or the other follows a bernoulli distribution (two discrete outcomes) and expected value is the same as the probability of a given outcome for a bernoulli distribution], and is shown in the derivation below.

$$
\begin{aligned}
B(\hat{p_0}) &= E(\hat{p_0}) - p_0 \\
B(\hat{p_0}) &= P(Y_i = 0) - P(Y_i = 0) \\
B(\hat{p_0}) &= 0
\end{aligned}
$$

(E, F) To get a sense of the bivariate normal distribution and whether it's suitable for our problem, we create the **rbinorm** function to randomly generate pairs (that could represent, for instance, length and width measurements), then observe how the contour of the generated points changes as the values of the function's paramaters are adjusted. The last of the figures below shows that the cloud of randomly generated points approaches an elliptical contour as the number of points generated increases.

```{r}

rbinorm <- function(n, mu1, mu2, rho, sigma1_sq, sigma2_sq) {
  # returns an n x 2 matrix
  # where each row is realization of the bivariate normal distribution
  
  sigma1 <- sqrt(sigma1_sq)
  sigma2 <- sqrt(sigma2_sq)
  
  u <- rnorm(n)
  v <- rnorm(n)
  
  x1 <- mu1 + sigma1 * sqrt(1 - rho^2) * u + sigma1 * rho * v
  x2 <- mu2 + sigma2 * v
  
  return(cbind(x1, x2))
}

# rho = -0.7
genpoints<-rbinorm(50, 1, 3, -0.7, 1, 2)
plot(genpoints, xlab="X1", ylab="X2")

# rho = 0.1
genpoints<-rbinorm(50, 1, 3, 0.1, 1, 2)
plot(genpoints, xlab="X1", ylab="X2")

# rho = 0.9
genpoints<-rbinorm(50, 1, 3, 0.9, 1, 2)
plot(genpoints, xlab="X1", ylab="X2")

# testing with more samples to show elliptically contoured shape
genpoints <- rbinorm(5000, 1, 3, 0.9, 1, 2)
plot(genpoints, xlab="X1", ylab="X2")

  
```

(G) [intro + explanation] [Verify that, as the number of realiza- tions gets larger, we approximate μ and Σ with increasing precision.]

```{r}

M <- rbinorm(50, 1, 3, 0.7, 1, 3)
colMeans(M)
cov(M)

N <- rbinorm(10000, 1, 3, 0.7, 1, 3)
colMeans(N)
cov(N)

```
(H, I) ...

```{r}
...
```

(J) ...

$$

\begin{aligned}

  p_0f_0(x) &> p_1f_1(x) \\
  \hat{p_0}(\frac{1}{2\pi\sqrt{\text{det}\hat{\Sigma}}_0})e^{-\frac{1}{2}r_0^2(x)} &> \hat{p_1}(\frac{1}{2\pi\sqrt{\text{det}\hat{\Sigma}}_1})e^{-\frac{1}{2}r_1^2(x)} \\
  (\frac{\hat{p_0}}{\sqrt{\text{det}\hat{\Sigma}}_0})e^{-\frac{1}{2}r_0^2(x)} &> (\frac{\hat{p_1}}{\sqrt{\text{det}\hat{\Sigma}}_1})e^{-\frac{1}{2}r_1^2(x)} \\
  \frac{\hat{p_0}}{\hat{p_1}}\frac{\sqrt{\text{det}\hat{\Sigma}_1}}{\sqrt{\text{det}\hat{\Sigma}_0}}\frac{e^{-\frac{1}{2}r^2_0(x)}}{e^{-\frac{1}{2}r^2_1(x)}} &> 1 \\
  \log{\frac{\hat{p_0}}{\hat{p_1}}\frac{\sqrt{\text{det}\hat{\Sigma}_1}}{\sqrt{\text{det}\hat{\Sigma}_0}}\frac{e^{-\frac{1}{2}r^2_0(x)}}{e^{-\frac{1}{2}r^2_1(x)}}} &> \log{1} \\
  \log{\frac{\hat{p_0}}{\hat{p_1}}} + \log{\frac{\sqrt{\text{det}\hat{\Sigma}_1}}{\sqrt{\text{det}\hat{\Sigma}_0}}} + \log{\frac{e^{-\frac{1}{2}r^2_0(x)}}{e^{-\frac{1}{2}r^2_1(x)}}} &> 0 \\
  \log{\frac{\hat{p_0}}{\hat{p_1}}} + \frac{1}{2}\log{\frac{\text{det}\hat{\Sigma}_1}{\text{det}\hat{\Sigma}_0}} + \frac{1}{2}(r^2_1(x) - r^2_0(x)) &> 0 \\
  2\log{\frac{\hat{p_0}}{\hat{p_1}}} + \log{\frac{\text{det}\hat{\Sigma}_1}{\text{det}\hat{\Sigma}_0}} + (r^2_1(x) - r^2_0(x)) &> 0 \\
  2\log{\frac{\hat{p_0}}{\hat{p_1}}} + \log{\frac{\text{det}\hat{\Sigma}_1}{\text{det}\hat{\Sigma}_0}} + r^2_1(x) &> r^2_0(x) \\

\end{aligned}

$$
(K) ...

```{r}
classify <- function(width, length) {
  # returns 0 if flower is of type 0
  # and 1 if flower is of type 1
  
  r0_sq <- (...)
  r1_sq <- (...)
  
  if (r0_sq < log(.../...) + 2*log(.../...) + r1_sq) {
    return(0)
  } else {
    return(1)
  }
}
```

[brief conclusion]