---
title: "Report"
output: html_document
---

```{r}
data(iris)
d<-iris

library(tidyverse)
# x0 <- d %>% filter(Species=="versicolor") %>% select(3:4)
# x1 <- d %>% filter(Species=="virginica") %>% select(3:4)

iris %>%
  filter(Species %in% c("versicolor", "virginica")) %>%
  ggplot(aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  geom_point()
```

If the flower has petal length of 4.0 and petal width of 1.1, then we would assume it is a versicolor. Likewise, if the length is 6 and its width is 2.3, we would assume it is a virginica. If its length is 5.0 and width is 1.7, with a K-nearest neighbor model with K=3, it would be classified as a versicolor. However, with a different K value, the classification may be different. 

```{r}

rbinorm <- function(n, mu1, mu2, rho, sigma1sq, sigma2sq) {
  # returns a n x 2 matrix, each row is 
  sigma1 <- sqrt(sigma1sq)
  sigma2 <- sqrt(sigma2sq)
  
  u <- rnorm(n)
  v <- rnorm(n)
  
  x1 <- mu1 + sigma1 * sqrt(1 - rho^2) * u + sigma1 * rho * v
  x2 <- mu2 + sigma2 * v
  
  return(cbind(x1, x2))
}

# rho = -0.7
genpoints<-rbinorm(50, 1, 3, -0.7, 1, 2)
plot(genpoints, xlab="X1", ylab="X2")

# rho = 0.1
genpoints<-rbinorm(50, 1, 3, 0.1, 1, 2)
plot(genpoints, xlab="X1", ylab="X2")

# rho = 0.9
genpoints<-rbinorm(50, 1, 3, 0.9, 1, 2)
plot(genpoints, xlab="X1", ylab="X2")

# testing with more samples to show elliptically contoured shape
genpoints <- rbinorm(5000, mu1=1, mu2=3, rho=0.9, sigma1sq=1, sigma2sq=2)
plot(genpoints, xlab="X1", ylab="X2")

  
```

```{r}
rbinorm <- function(n, mu1, mu2, rho, sigma1sq, sigma2sq) {
  # returns a n x 2 matrix, each row is 
  sigma1 <- sqrt(sigma1sq)
  sigma2 <- sqrt(sigma2sq)
  
  u <- rnorm(n)
  v <- rnorm(n)
  
  x1 <- mu1 + sigma1 * sqrt(1 - rho^2) * u + sigma1 * rho * v
  x2 <- mu2 + sigma2 * v
  
  return(cbind(x1, x2))
}

M <- rbinorm(50, mu1=1, mu2=3, rho=0.7, sigma1sq=1, sigma2sq=3)
colMeans(M)
cov(M)

N <- rbinorm(10000, mu1=1, mu2=3, rho=0.7, sigma1sq=1, sigma2sq=3)
colMeans(N)
cov(N)

```

Verify that, as the number of realiza- tions gets larger, we approximate μ and Σ with increasing precision.

```{r}
rbinorm <- function(n, mu1, mu2, rho, sigma1sq, sigma2sq) {
  # returns a n x 2 matrix, each row is 
  sigma1 <- sqrt(sigma1sq)
  sigma2 <- sqrt(sigma2sq)
  
  u <- rnorm(n)
  v <- rnorm(n)
  
  x1 <- mu1 + sigma1 * sqrt(1 - rho^2) * u + sigma1 * rho * v
  x2 <- mu2 + sigma2 * v
  
  return(cbind(x1, x2))
}

x0 <- d %>% filter(Species=="versicolor") %>% select(3:4)
x1 <- d %>% filter(Species=="virginica") %>% select(3:4)


```


